---
title: "FFTrees"
bibliography: homepage.bib
---

The `FFTrees` package creates fast and frugal decision trees (FFTs). Fast and frugal decision trees are very simple [heuristics](https://en.wikipedia.org/wiki/Heuristic) decision algorithms that, in contrast to regression and more complex algorithms such as [random forests](https://en.wikipedia.org/wiki/Random_forest) and (non-frugal) [decision trees](https://en.wikipedia.org/wiki/Decision_tree), allow people to make fast decisions based on minimal information. FFTs are not just easy to use, they can also predict data as well, if not better, than more complex algorithms such as regression [@gigerenzer1999simple;@martignon2008categorization]. Because they are so effective and easy to use and, fast and frugal trees have been used to improve decision making in medical [@green1997alters; @jenny2013simple] and financial [@neth2014homo] contexts.

The purpose of `FFTrees` is to allow anyone to quickly and easily generate, visualise, and use FFTs for their own data.

```{r eval = FALSE}
# Install FFTrees from CRAN (only necessary once)
install.packages("FFTrees")
```



```{r, fig.width = 7, fig.height = 6.5, message = FALSE, warning = FALSE}
library(FFTrees)

heart.fft <- FFTrees(formula = diagnosis ~.,        # Criterion
                     data = heart.train,            # Training data   
                     data.test = heart.test)        # Testing data

plot(heart.fft,                                     # FFTrees object
     main = "Heart Disease",                        # Plot title
     decision.names = c("Healthy", "Diseased"),     # Decision names
     data = "test")                                 # Show statistics for test data
```

## Manuscript

You can read more about the FFTrees package in our accompanying manuscript [FFTrees: A toolbox to create, visualize, and evaluate fast-and-frugal decision trees](http://journal.sjdm.org/17/17217/jdm17217.pdf) [@phillips2017fftrees], co-authored with [Hansjörg Neth](https://www.spds.uni-konstanz.de/hans-neth), [Jan Woike](https://www.mpib-berlin.mpg.de/en/staff/jan-k-woike) and [Wolfgang Gaissmaier](https://www.spds.uni-konstanz.de/wolfgang-gaissmaier)

## ShinyFFTrees

- [ShinyFFTrees Link](https://econpsychbasel.shinyapps.io/ShinyFFTrees/)

Through the magic of the internet (and the amazing Shiny framework), you can even play around with FFTrees entirely in a web-browser. Either click the link above, or play around in the window below:

<iframe src="https://econpsychbasel.shinyapps.io/ShinyFFTrees/", height = 500px, width = 900px, align = middle></iframe>


# References
