---
title: "R Packages"
---

Here is a list of my R packages available either on CRAN or GitHub:

## yarrr

The `yarrr` package contains a mixture of functions and datasets that support my e-book YaRrr! The pirate's guide to R. The most notable function in the package is `pirateplot()`, a modification of the fantastic beanplot from the `beanplot` package that shows the relationship between a categorical independent variable and a numeric dependent variable. The plot replaces outdated barplots by showing Raw data, Descriptive statistics, and Inferential statistics.



```{r, echo = FALSE, fig.width = 6, fig.height = 5, message = FALSE, warning = FALSE, fig.align='center'}
library(yarrr)
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme = 1,
           back.col = "white",
           gl.col = "white",
           bean.f.o = c(0, .1, .7, .1),
        #   bean.b.o = c(0, .1, 1, .1),
           point.o = c(.4, .1, .1, .1),
           avg.line.o = c(.3, 1, .3, .3),
           inf.f.o = c(.1, .1, .1, .9),
           bar.f.o = c(.1, .8, .1, .1),
           inf.f.col = c("white", "white", "white", piratepal("xmen")[4]),
           main = "4 Elements of a pirateplot", 
           pal = "xmen")


text(.7, 350, labels = "Points")
text(.7, 345, labels = "Raw Data", pos = 1, cex = .8)
arrows(.7, 310, .97, 270, length = .1)

text(1.4, 200, labels = "Bar/Line")
text(1.4, 200, labels = "Center", pos = 1, cex = .8)
arrows(1.4, 170, 1.54, 125, length = .1)

text(2.4, 250, labels = "Bean")
text(2.4, 250, labels = "Density", pos = 1, cex = .8)
arrows(2.4, 220, 2.85, 200, length = .1)

text(3.55, 300, labels = "Band")
text(3.55, 290, labels = "Inference\n95% HDI or CI", pos = 1, cex = .8)

arrows(3.55, 240, 3.8, 150, length = .1)
```

You can install the most stable version of `yarrr` from CRAN, or the latest developer version from GitHub as follows:

```{r eval = FALSE}
install.packages("yarrr")
# devtools::install_github("ndphillips/yarrr", build_vignettes = TRUE) # developer version
```

Here is an example of a pirateplot showing the relationship between the age and weight of chickens:

```{r fig.width = 8, fig.height = 6, fig.align='center'}
# Create a pirateplot from the ChickWeight data
yarrr::pirateplot(formula = weight ~ Time,
                  data = ChickWeight,
                  main = "A Pirateplot of chicken weights by time")
```

As you can see, the plot includes all the raw data, 95% Bayesian Highest Density Intervals (HDIs), smoothed densities, and bright colors by default. The `pirateplot()` function contains many arguments that allow you to customise the look of the plot. 

## FFTrees

The `FFTrees` package creates fast and frugal decision trees (FFTs). Fast and frugal decision trees are decision algorithms 

```{r eval = FALSE}
install.packages("FFTrees")
# devtools::install_github("ndphillips/FFTrees", build_vignettes = TRUE) # developer version
```



```{r, fig.width = 8, fig.height = 7, message = FALSE, warning = FALSE}
library(FFTrees)

set.seed(100) # For cross validation replication
heart.fft <- FFTrees(formula = diagnosis ~.,        # formula
                     data = heartdisease,           # dataframe
                     train.p = .5)                  # cross validation split

plot(heart.fft,                                     # FFTrees object
     main = "Heart Disease",                        # Plot title
     decision.names = c("Healthy", "Diseased"),     # Decision names
     data = "test")                                 # Show statistics for test data
```


