---
title: "R Packages"
bibliography: homepage.bib
---

Here is a list of my R packages available either on CRAN or GitHub:

# yarrr

The `yarrr` package contains a mixture of functions and datasets that support my e-book [YaRrr! The pirate's guide to R](http://www.thepiratesguidetor.com). The most notable function in the package is `pirateplot()`. A pirateplot shows the relationship between a categorical independent variable and a numeric dependent variable. However, in contrast to notoriously deceptive barplots, a pirateplot reveals patterns in data by showing Raw data (as points), Descriptive statistics (as lines and/or bars), and Inferential statistics (as bands).



```{r, echo = FALSE, fig.width = 6, fig.height = 5, message = FALSE, warning = FALSE, fig.align='center'}
library(yarrr)
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme = 1,
           back.col = "white",
           gl.col = "white",
           bean.f.o = c(0, .1, .7, .1),
        #   bean.b.o = c(0, .1, 1, .1),
           point.o = c(.4, .1, .1, .1),
           avg.line.o = c(.3, 1, .3, .3),
           inf.f.o = c(.1, .1, .1, .9),
           bar.f.o = c(.1, .8, .1, .1),
           inf.f.col = c("white", "white", "white", piratepal("xmen")[4]),
           main = "4 Elements of a pirateplot", 
           pal = "xmen")


text(.7, 350, labels = "Points")
text(.7, 345, labels = "Raw Data", pos = 1, cex = .8)
arrows(.7, 310, .97, 270, length = .1)

text(1.4, 200, labels = "Bar/Line")
text(1.4, 200, labels = "Center", pos = 1, cex = .8)
arrows(1.4, 170, 1.54, 125, length = .1)

text(2.4, 250, labels = "Bean")
text(2.4, 250, labels = "Density", pos = 1, cex = .8)
arrows(2.4, 220, 2.85, 200, length = .1)

text(3.55, 300, labels = "Band")
text(3.55, 290, labels = "Inference\n95% HDI or CI", pos = 1, cex = .8)

arrows(3.55, 240, 3.8, 150, length = .1)
```

You can install the most stable version of `yarrr` from CRAN, or the latest developer version from GitHub as follows:

```{r eval = FALSE}
install.packages("yarrr")
# devtools::install_github("ndphillips/yarrr", build_vignettes = TRUE) # developer version
```

Here is an example of a pirateplot showing the relationship between the age and weight of chickens:

```{r fig.width = 8, fig.height = 6, fig.align='center'}
# Create a pirateplot from the ChickWeight data
yarrr::pirateplot(formula = weight ~ Time,
                  data = ChickWeight,
                  main = "A Pirateplot of chicken weights by time")
```

The `pirateplot()` function contains many arguments that allow you to easily customise the look of the plot, from the colors, to the element transparencies. To learn more, look at the `pirateplot()` vignette [on CRAN](https://cran.r-project.org/web/packages/yarrr/vignettes/pirateplot.html).

# FFTrees

The `FFTrees` package creates fast and frugal decision trees (FFTs). Fast and frugal decision trees are very simple [heuristics](https://en.wikipedia.org/wiki/Heuristic) decision algorithms that, in contrast to regression and more complex algorithms such as [random forests](https://en.wikipedia.org/wiki/Random_forest) and (non-frugal) [decision trees](https://en.wikipedia.org/wiki/Decision_tree), allow people to make fast decisions based on minimal information. FFTs are not just easy to use, they can also predict data as well, if not better, than more complex algorithms such as regression [@gigerenzer1999simple;@martignon2008categorization]. Because they are so effective and easy to use and, fast and frugal trees have been used to improve decision making in medical [@green1997alters; @jenny2013simple] and financial [@neth2014homo] contexts.

The purpose of `FFTrees` is to allow anyone to quickly and easily generate, visualise, and use FFTs for their own data.

```{r eval = FALSE}
install.packages("FFTrees")
# devtools::install_github("ndphillips/FFTrees", build_vignettes = TRUE) # developer version
```



```{r, fig.width = 8, fig.height = 7, message = FALSE, warning = FALSE}
library(FFTrees)

set.seed(100) # For cross validation replication
heart.fft <- FFTrees(formula = diagnosis ~.,        # formula
                     data = heartdisease,           # dataframe
                     train.p = .5)                  # cross validation split

plot(heart.fft,                                     # FFTrees object
     main = "Heart Disease",                        # Plot title
     decision.names = c("Healthy", "Diseased"),     # Decision names
     data = "test")                                 # Show statistics for test data
```

# References
